Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}

### 2023-05-09 Papers/Paper CASC format/MLaTeX/CLpaper.tex
...

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[latin1]{inputenc}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{ifthen}
\usepackage{mmacells}
\usepackage[invisible]{MLaTeX}

@@

\begin{document}
%
\title{Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
%
%\titlerunning{Abbreviated paper title}
%
\author{Diogo A. Gomes\inst{1}\orcidID{0000-0002-3129-3956} \and
	Friedemann Krannich\inst{1}\orcidID{1111-2222-3333-4444} \and
	Bashayer Majrashi\inst{1}\orcidID{0009-0008-3677-8064 } \and
	Ricardo de Lima Ribeiro\inst{1}\orcidID{0000-0002-2707-2210}}
%
\authorrunning{D.A. Gomes et al.}
%
\institute{King Abdullah University of Science and Technology (KAUST), AMCS/CEMSE Division, Thuwal 23955-6900, Saudi Arabia\
	\email{{diogo.gomes,friedemann.krannich,bashayer.majrashi, ricardo.ribeiro}@kaust.edu.sa}}
%
\maketitle
%
\begin{abstract}
	Many partial differential equations (PDEs) admit conserved quantities like mass or energy. These quantities are often essential for establishing well-posedness results. When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved over time by the scheme. Such conservation may ensure the stability of the numerical scheme. We present an algorithm for verifying the preservation of a polynomial quantity under a polynomial finite-difference scheme. Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables. Additionally, we introduce an algorithm for finding conserved quantities, given a scheme. We illustrate our algorithm with several finite-difference schemes. Our approach incorporates parametric G\"obner bases to handle parameters, ensuring accurate computation of conserved quantities.

@@

\section{Introduction}
%\label{intro}
{\color{blue} add 
Comprehensive Gr{\"o}bner Systems 
}
In many partial differential equations (PDEs), conserved integral quantities play a crucial role, as they often represent properties of physical significance, such as energy or mass. For example, the advection equation preserves energy and the heat equation conserves mass. 
Since closed-form solutions for these PDEs are often unavailable, finite-difference schemes serve as a practical numerical alternative. Hence, it becomes imperative to determine whether the chosen scheme preserves these conserved quantities. This information is not only valuable for gauging the accuracy of the approximation, but also vital for evaluating its stability. Automating conservation computations in computer algebra systems streamlines the process, saves time, enhances accuracy, and may create new research opportunities.

@@

Here, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. %We restrict $G$ to the set of polynomials in $k$ variables, because our algorithm below can only treat polynomial quantities.
Let $\tilde{n}:=(n_1,\dots,n_d,s_1,\dots,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
	\mathcal{J}[u](t_1,\ldots,t_{l})
	=\sum_{\tilde{n}\in\Ii\times\Nn^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
	%=\sum_{n\in\Ii}G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))\delta_{t_1,\ldots,t_{d-l+1}}(n_l,\ldots,n_d)
Let $u,v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, then
	D^l\mathcal{J}[u](v)(t_1,\ldots,t_{l}):=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\Big|_{\epsilon=0}(t_1,\ldots,t_{l})\\
	=\sum_{\tilde{n}}D_1G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^1)\\
	+D_2G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^2)\\+\ldots
	+D_kG(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^k)\\
	=\sum_{\tilde{n}}D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1,\ldots,t_{l})}(s_1-e^1_{d+1},\ldots,s_l-e^1_{d+l})v(\tilde{n})\\
	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1,\ldots,t_{l})}(s_1-e^2_{d+1},\ldots,s_l-e^2_{d+l})v(\tilde{n})\\+\ldots
	+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1,\ldots,t_{l})}(s_1-e^k_{d+1},\ldots,s_l-e^k_{d+l})v(\tilde{n})\\
	=\sum_{\tilde{n}}\Big(D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1+e^1_{d+1},\ldots,t_{l}+e^1_{d+l})}(s_1,\ldots,s_l)\\
	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1+e^2_{d+1},\ldots,t_{l}+e^2_{d+l})}(s_1,\ldots,s_l)\\+\ldots
	+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1+e^k_{d+1},\ldots,t_{l}+e^k_{d+l})}(s_1,\ldots,s_l)\Big)v(\tilde{n})\\
	=:\sum_{\tilde{n}}\mathcal{V}^l(\mathcal{J})[u](\tilde{n},t)v(\tilde{n}).
$\mathcal{V}^l(\mathcal{J})$ is the \emph{discrete partial variational derivative} of $\mathcal{J}$.
	Note that $\mathcal{V}^0=\mathcal{V}$.
	\label{dpvd1}
	The discrete partial variational derivative of 
		\mathcal{J}[u](t)=\sum_nu(n,t+1)-u(n,t)=\sum_{n,s}(u(n,s+1)-u(n,s))\delta_t(s)
	is $\delta_{t+1}(s)-\delta_t(s)$, because
	\small
		D^1(\mathcal{J})[u](v)(t)\\
		=\frac{d}{d\epsilon} \sum_{n,s}((u(n,s+1)+\epsilon v(n,s+1))-(u(n,s)+\epsilon v(n,s)))\delta_t(s)\\
		=\sum_{n,s}v(n,s+1)\delta_t(s)-v(n,s)\delta_t(s)=\sum_{n,s}v(n,s)\delta_t(s-1)-v(n,s)\delta_t(s)\\
		=\sum_{n,s}(\delta_{t+1}(s)-\delta_t(s))v(n,s)=\sum_{n,s}\mathcal{V}^1(\mathcal{J})(n,s,t)v(n,s).
	In our {\sc Mathematica} implementation, Example \ref{dpvd1} is computed by
	\begin{EXE}
		LABEL="dpvd1"
		(*CODE*)
		variables = Association["indVars"->{n,t},"depVars"->{u},"timeVars"->{t}]
		expression = u[n,t+1] - u[n,t]
		PartialDVarDOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}	
	\begin{small}
		\input{executables/dpvd1ctt.ed.tex}
		\input{executables/dpvd1ctt.result.ed.tex}
	\end{small}
We have a result similar to Theorem \ref{theo1} for the discrete partial variational derivative:
\begin{theorem}\label{theo2}
	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. If $\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
	$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$.
	Conversely,
	$\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
\end{theorem}
The proof is similar to the proof of Theorem \ref{theo1}.

@@

 \newpage 
%just for now 
\section{Parametric Groebner Systems}
\label{pgs}

% This section covers key concepts for studying parametric 
% polynomial ideals, including parametric rings and specialization 
% ideals and comprehensive Gr{\"o}bner systems for parametric polynomials that generalize 
% Gr{\'o}bner basis in section \ref{4sec1}. Section 2 explains why comprehensive Gr{\"o}bner systems are advantageous 
% and some problems that arise without CGS. Section 3 
% extends the ideal membership problem to parametric ideals.
In this section, we delve into essential concepts for 
parametric polynomial ideals. These include parametric rings, 
specialized ideals, and comprehensive Gr{\"o}bner systems, which are built upon 
Gr{\"o}bner basis theory. 
We discuss the benefits of using comprehensive Gr{\"o}bner systems 
and how they help address problems that occur when not using them. 
Furthermore, we expand on the ideal membership problem to encompass parametric ideals.

% \subsection{Parametric Polynomials and Comprehensive Groebner Systems}
% \label{4sec1}

% This section introduces the general setting of parametric polynomial rings and a ring homomorphism that 
% represents the choice of parameters. Moreover, we define comprehensive Gr{\"o}bner systems (CGS),
% and show some examples of (CGS) using our implementation of a parametric Buchberger's algorithm.

@@

\end{definition}
% For example let $P=a x+by \in \Rr[a,b][x,y]$. 
% a specialization $\sigma$ is defined by $P\rightarrow P(2,5)$ so that $$\sigma (P)=\sigma(a x+by)$$
% and since $\sigma $ is a ring homomorphism we have,
% $$\sigma(a x+by)= \sigma(a) \sigma(x)+\sigma(b)\sigma(y)$$
% $$= 2x+5y.$$
It is easier to think about specializations as a choice of parameters. \\

Comprehensive Gr{\"o}bner systems extend Gr{\"o}bner basis to 
parametric polynomials and have diverse practical applications in 
algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.
\begin{definition}
    Let $\Ii$ be a finitely generated, nonzero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
    Let $\{A_1,\dots,A_k\}$ be a family of locally closed sets that forms a partition of $L^m$ 
    , such that the disjoint union $\bigcup\limits_{i=1}^k A_i=L^m$, and $\{G_1,\dots,G_k\}$ 
    be a family of subsets in $\mathcal{K}[A][X]$.
    The system, 
    \[   \left\{
    \begin{array}{ll}
        G_1 & A_1 \\
        G_2  & A_2\\
        \vdots  & \vdots \\
        G_k & A_k \\
    \end{array} 
    \right. \]

@@

A proof of existence and the first algorithm to compute the (CGS) were introduced by Volker Weispfenning.
\begin{theorem}
    Let $\mathcal{I}\subseteq\mathcal{K}[A][X].$ The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
\end{theorem}
We use our implementation of a parametric Buchberger's algorithm to compute (CGS). 
This is illustrated in the following examples.
\begin{example}
    Let $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by 
    $$\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.$$
    The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
    {\color{red} we use our code here}
	% $$
    % % \begin{cases}\left\{a^2 b^2 y^3+a b^4 y^4-2 a b^3 y^3,x-b^2
    % %     y^2\right\} &  a b \neq 0 \\
    % % \{x\} &  b=0 \\
    % % \left\{b^2 y^2-x\right\} &  a=0 \land b \neq 0.
    % % \end{cases}
    % $$
\end{example}
The idea of a comprehensive Gr{\"o}bner system is to provide a "parameter sensitive Gr{\"o}bner basis".
It branches the cases of the parameters that affect leading terms. For instance, in the previous
example if $a=0$ then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
the (CGS) considers cases of the leading coefficients parameters Vanishing.

@@

Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
In practice, to compute the Gr{\"o}bner basis, we assume non-zero parameters by default.
This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialization ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
 The impact of this assumption is seen in (i) and (ii):
\begin{enumerate}
    \item In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
    \begin{example}
        \label{ex431}
        Let $\Ii=\langle 
           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis 
           is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
           the specialization $\sigma \in \Rr^2$ given by $P\rightarrow P(0,1)$.
           Now, We claim that the set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
           To show that we find the S-polynomials of each pair in the set and we find that: $S(-y^2, -2 + x y)=-2 y$.
           We divide $S(-y^2, -2 + x y)$ by $\sigma(\Gg)$ and find that the remainder is $-2y$. 
           Therefore by the Buchberger Criterion, $\sigma(\Gg)$ is not a Gr{\"o}bner basis.
    \end{example}
    \item Consequently, the algorithm utilized to resolve the ideal membership 
    problem in \ref{sec2.1} can not be used for all specializations.

    \begin{example}
        \label{ex433}
        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero parameters
        is $$\sigma_0(\Gg)=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$$ where $\sigma_0 \in \Rr^2\backslash V(a,b)$. 
        Now, it is possible to check if a polynomial $f\in \Rr[x,y]$ belongs to $\sigma(\Ii)$ only for $\sigma_0 $     
        because we know that $\sigma_0(\Gg)$ is a Gr{\"o}bner basis. For other specializations such as 
        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
        Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used $\sigma(\Gg)$.

@@

@@

\noindent{\bf Step 1: Build the discrete time derivative.}
We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is performed in our code using {\sc TimeDifference}.
\begin{example}
	The discrete time derivative for $\sum_n u(n,t)$ is
	\begin{EXE}
		LABEL="heateq2"
		(*CODE*)
		variables = <|"indVars"->{n,t},"depVars"->{u}|>
		expression = u[n,t]
		TimeDifferenceOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}

	\begin{small}
		\input{executables/heateq2ctt.ed.tex}
		\input{executables/heateq2ctt.result.ed.tex}
	\end{small}
\end{example}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{graphicx}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[latin1]{inputenc}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{ifthen}
\usepackage{mmacells}

@@

\begin{document}
\title{Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
\author{Diogo A. Gomes\inst{1}\orcidID{0000-0002-3129-3956} \and
	Friedemann Krannich\inst{1}\orcidID{1111-2222-3333-4444} \and
	Bashayer Majrashi\inst{1}\orcidID{0009-0008-3677-8064 } \and
	Ricardo de Lima Ribeiro\inst{1}\orcidID{0000-0002-2707-2210}}
\authorrunning{D.A. Gomes et al.}
\institute{King Abdullah University of Science and Technology (KAUST), AMCS/CEMSE Division, Thuwal 23955-6900, Saudi Arabia\
	\email{{diogo.gomes,friedemann.krannich,bashayer.majrashi, ricardo.ribeiro}@kaust.edu.sa}}
\maketitle
\begin{abstract}
	Many partial differential equations (PDEs) admit conserved quantities like mass or energy. These quantities are often essential for establishing well-posedness results. When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved over time by the scheme. Such conservation may ensure the stability of the numerical scheme. We present an algorithm for verifying the preservation of a polynomial quantity under a polynomial finite-difference scheme. Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables. Additionally, we introduce an algorithm for finding conserved quantities, given a scheme. We illustrate our algorithm with several finite-difference schemes. Our approach incorporates parametric G\"obner bases to handle parameters, ensuring accurate computation of conserved quantities.

@@

\section{Introduction}
{\color{blue} add 
Comprehensive Gr{\"o}bner Systems 2222
}
In many partial differential equations (PDEs), conserved integral quantities play a crucial role, as they often represent properties of physical significance, such as energy or mass. For example, the advection equation preserves energy and the heat equation conserves mass. 
Since closed-form solutions for these PDEs are often unavailable, finite-difference schemes serve as a practical numerical alternative. Hence, it becomes imperative to determine whether the chosen scheme preserves these conserved quantities. This information is not only valuable for gauging the accuracy of the approximation, but also vital for evaluating its stability. Automating conservation computations in computer algebra systems streamlines the process, saves time, enhances accuracy, and may create new research opportunities.

@@

Here, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. 
Let $\tilde{n}:=(n_1,\dots,n_d,s_1,\dots,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
\begin{gather*}
	\mathcal{J}[u](t_1,\ldots,t_{l})
	=\sum_{\tilde{n}\in\Ii\times\Nn^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
\end{gather*}
Let $u,v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, then
\footnotesize
\begin{gather*}
	D^l\mathcal{J}[u](v)(t_1,\ldots,t_{l}):=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\Big|_{\epsilon=0}(t_1,\ldots,t_{l})\\
	=\sum_{\tilde{n}}D_1G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^1)\\
	+D_2G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^2)\\+\ldots
	+D_kG(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^k)\\
	=\sum_{\tilde{n}}D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1,\ldots,t_{l})}(s_1-e^1_{d+1},\ldots,s_l-e^1_{d+l})v(\tilde{n})\\
	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1,\ldots,t_{l})}(s_1-e^2_{d+1},\ldots,s_l-e^2_{d+l})v(\tilde{n})\\+\ldots
	+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1,\ldots,t_{l})}(s_1-e^k_{d+1},\ldots,s_l-e^k_{d+l})v(\tilde{n})\\
	=\sum_{\tilde{n}}\Big(D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1+e^1_{d+1},\ldots,t_{l}+e^1_{d+l})}(s_1,\ldots,s_l)\\
	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1+e^2_{d+1},\ldots,t_{l}+e^2_{d+l})}(s_1,\ldots,s_l)\\+\ldots
	+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1+e^k_{d+1},\ldots,t_{l}+e^k_{d+l})}(s_1,\ldots,s_l)\Big)v(\tilde{n})\\
	=:\sum_{\tilde{n}}\mathcal{V}^l(\mathcal{J})[u](\tilde{n},t)v(\tilde{n}).
\end{gather*}
\normalsize
\begin{definition}
$\mathcal{V}^l(\mathcal{J})$ is the \emph{discrete partial variational derivative} of $\mathcal{J}$.
	Note that $\mathcal{V}^0=\mathcal{V}$.
\end{definition}
\begin{example}
	\label{dpvd1}
	The discrete partial variational derivative of 
	\begin{gather*}
		\mathcal{J}[u](t)=\sum_nu(n,t+1)-u(n,t)=\sum_{n,s}(u(n,s+1)-u(n,s))\delta_t(s)
	\end{gather*}
	is $\delta_{t+1}(s)-\delta_t(s)$, because
	\small
	\begin{gather*}
		D^1(\mathcal{J})[u](v)(t)\\
		=\frac{d}{d\epsilon} \sum_{n,s}((u(n,s+1)+\epsilon v(n,s+1))-(u(n,s)+\epsilon v(n,s)))\delta_t(s)\\
		=\sum_{n,s}v(n,s+1)\delta_t(s)-v(n,s)\delta_t(s)=\sum_{n,s}v(n,s)\delta_t(s-1)-v(n,s)\delta_t(s)\\
		=\sum_{n,s}(\delta_{t+1}(s)-\delta_t(s))v(n,s)=\sum_{n,s}\mathcal{V}^1(\mathcal{J})(n,s,t)v(n,s).
	\end{gather*}
	\normalsize
\end{example}
\begin{example}
	In our {\sc Mathematica} implementation, Example \ref{dpvd1} is computed by

	\begin{small}
		\begin{mmaCell}[moredefined={variables, expression, PartialDVarD}]{Input}
  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\},
  "timeVars"\(\pmb{\to}\)\{t\}|>
  PartialDVarD[variables][u[n,t+1]-u[n,t]]

\end{mmaCell}
		\begin{mmaCell}{Output}
  \{-KroneckerDelta[t]+KroneckerDelta[1+t]\}
\end{mmaCell}
	\end{small}
\end{example}
We have a result similar to Theorem \ref{theo1} for the discrete partial variational derivative:
\begin{theorem}\label{theo2}
	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. If $\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
	$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$.
	Conversely,
	$\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
\end{theorem}
The proof is similar to the proof of Theorem \ref{theo1}.

@@

 \newpage 
\section{Parametric Groebner Systems}
\label{pgs}

In this section, we delve into essential concepts for 
parametric polynomial ideals. These include parametric rings, 
specialized ideals, and comprehensive Gr{\"o}bner systems, which are built upon 
Gr{\"o}bner basis theory. 
We discuss the benefits of using comprehensive Gr{\"o}bner systems 
and how they help address problems that occur when not using them. 
Furthermore, we expand on the ideal membership problem to encompass parametric ideals.

@@

\end{definition}
It is easier to think about specializations as a choice of parameters. \\

Comprehensive Gr{\"o}bner systems extend Gr{\"o}bner basis to 
parametric polynomials and have diverse practical applications in 
algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.
\begin{definition}
    Let $\Ii$ be a finitely generated, nonzero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
    Let $\{A_1,\dots,A_k\}$ be a family of locally closed sets that forms a partition of $L^m$ 
    , such that the disjoint union $\bigcup\limits_{i=1}^k A_i=L^m$, and $\{G_1,\dots,G_k\}$ 
    be a family of subsets in $\mathcal{K}[A][X]$.
    The system, 
	\begin{equation}\label{CGS}
	\left\{
		\begin{array}{ll}
			G_1 & A_1 \\
			\vdots  & \vdots \\ 
			G_k & A_k \\
		\end{array}
		\right. 
	\end{equation}

@@

A proof of existence and the first algorithm to compute the (CGS) were introduced by Volker Weispfenning.
\begin{theorem}
    Let $\mathcal{I}\subseteq\mathcal{K}[A][X].$ The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
\end{theorem}
We use our implementation of a parametric Buchberger's algorithm to compute (CGS). 
This is illustrated in the following examples.
\begin{example}
    Let $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by 
    $$\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.$$
    The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
    {\color{red} we use our code here}
\end{example}
The idea of a comprehensive Gr{\"o}bner system is to provide a "parameter sensitive Gr{\"o}bner basis".
It branches the cases of the parameters that affect leading terms. For instance, in the previous
example if $a=0$ then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
the (CGS) considers cases of the leading coefficients parameters Vanishing.

@@

Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
In practice, to compute the Gr{\"o}bner basis without considering the parameters, we assume non-zero parameters by default.
This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialized ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
Next, we show the impact of this assumption.
\begin{enumerate}
    \item In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
    \begin{example}
        \label{ex431}
        Let $\Ii=\langle 
           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis 
           is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
           the specialization $\sigma$ given by $P\rightarrow P(0,1)$.
           The set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
		   This can be shown by the Buchberger Criterion (see chapter 2 in\cite{Cox}).

    \end{example}
    \item Consequently, the algorithm utilized to resolve the ideal membership 
    problem in chapter 2 of \cite{Cox} can not be used for all specializations.

    \begin{example}
        \label{ex433}
        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero parameters
        is $$\sigma_0(\Gg)=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$$ where $\sigma_0 \in \Rr^2\backslash V(a,b)$. 
        Now, it is possible to check if a polynomial $f\in \Rr[x,y]$ belongs to $\sigma(\Ii)$ only for $\sigma_0 $     
        because we know that $\sigma_0(\Gg)$ is a Gr{\"o}bner basis. For other specializations such as 
        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
        Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used $\sigma(\Gg)$.  
    \end{example}

@@

Using CGS can resolve these problems. Problem 1 is fixed by the definition of (CGS), as we have $\sigma_i(G_i)$ a 
 Gröbner basis for any specialization $\sigma_i$. Moreover, 
 to discuses problem 2 we need to look at the parametric ideal membership first.

{\color{red} Is this a definition?}
 Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ under $\sigma$, if there exists a specialization 
 $\sigma\in L^m$ such that $\sigma(f)\in\sigma(\Ii)$.\\

To determine if $f\in\Ii$ we consider the (CGS) of $\Ii$ given \ref{CGS}. 
We check if $f\in\sigma_i(\Ii)$ for each $i\in\{1,..,k\}$ and $\sigma_i\in A_i$.
For each $\sigma_i(\Ii)$ we have the corresponding Gröbner basis $\sigma_i(G_i)$
so we divide $f$ by the $\sigma_i(G_i)$ if the remainder is zero for some $i\in\{1,..,k\}$,
then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$ we conclude that 
$f\notin\Ii$.

\begin{example}
	Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
    $$\Ff=\{a^2 x - y, -b^2 x y^2 - x\}.$$
	We will determine if $f(x,y)=-x+y^2\in\mathcal{I}$.
	We compute the corresponding (CGS)
	{\color{red} add example 3}\\
	First let $A_1=\{(a,b)~|~ a\ne 0, b\ne 0\}$ and $A_2=\{(a,b)~|~ ab=0\}$. 
	We divide $f$ by $\{\frac{y}{b^2}+y^3,x-\frac{y}{a^2}\}$ and we obtain the remainder $-\frac{y}{a^2}+y^2$ and 
	since the exists no specialization in $A_1$ such that this remainder is zero we conclude that $f\notin \sigma_1(\Ii)$ for
	$\sigma_1\in A_1$. Now we repeat the reduction of $f$ but with the second Gröbner basis $\{x,y\}$. We obtain a zero remainder, which imply that 
	$f\in \sigma_2(\Ii)$ for $\sigma_2\in A_2$. Therefore, $f\in\Ii$ under a specialization $\sigma_2\in A_2$.
\end{example}

@@

\noindent{\bf Step 1: Build the discrete time derivative.}
We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is performed in our code using {\sc TimeDifference}.
\begin{example}
	The discrete time derivative for $\sum_n u(n,t)$ is

	\begin{small}
		\begin{mmaCell}[moredefined={variables, expression, TimeDifference}]{Input}
  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\}|>
  TimeDifference[variables][u[n,t]]

\end{mmaCell}
		\begin{mmaCell}{Output}
  <|exp\(\to\)-u[n,t]+u[n,1+t]|>
\end{mmaCell}
	\end{small}
\end{example}
